GPU-based Adaptive Beamforming for Active Sonar Imaging

### 2012-05-07 ECUA2012/gpu_capon_ecua2012.tex
la2pdf commit - 2012:05:07	20:10:32	(Mon 7.May)

In a typical active sonar imaging scenario encoded signals are transmitted towards a surface of interest, and the resulting reflections are recorded using an array of receivers. The receiver outputs can then be coherently combined in the post-processing to focus on specific points in the image. The combination process involves assigning suitable delays and weights to the array's channels, and is commonly referred to as beamforming.

@@

The Minimum Variance Distortionless Response (MVDR, or Capon) beamformer is one such technique. While ensuring unit gain in the look direction, it computes the set of weights that minimizes the power accumulated by the array from other directions. This ability to suppress interference power improves the image detail level and \gls{SNR}, and the effect can be dramatic if the noise and interference is primarily focused around specific angles. The feasibility of using the \gls{MVDR} beamformer in active sonar imaging has been demonstrated in e.g.~\cite{Blomberg2012a,Blomberg2011,Dursun2009,Lo2004}.

However, despite its inherent potential, the \gls{MVDR} beamformer has yet to see widespread adoption in the active sonar community. There may be several reasons for this. First, the method is not inherently robust, and will suffer from a phenomena called signal cancellation in active systems. Second, in its original form, the computational complexity is cubic with the number of channels, O(M$^3$), while \gls{DAS} is only at O($M$). This makes it impractical in many situations, and even infeasible in high channel systems~\cite{So2011}. \todo{mention disadvantage of tuning parameters?}

We will show that the suggested means for making the \gls{MVDR} behave well in active systems will also make it less computationally complex. Furthermore, by implementing it on a modern \gls{GPU} we have gained a speedup of around 2 orders of magnitude compared to an optimized single thread implementations in C.

@@

@@

>>>>

In a typical active sonar imaging scenario encoded signals are transmitted towards a surface of interest, and the resulting reflections are recorded using an array of receivers. These receivers may be independently delayed and weighted to adjust the array's point of focus, a concept commonly referred to as beamforming.

@@

The Minimum Variance Distortionless Response (MVDR, or Capon) beamformer is one such technique. While ensuring unit gain in the look direction, it computes the set of weights that minimizes the power accumulated by the array from other directions. Its ability to suppress interference power and improve image quality in active sonar systems has been demonstrated in e.g.~\cite{Blomberg2012a,Blomberg2011,Dursun2009,Lo2004}.

However, despite its inherent potential, the \gls{MVDR} beamformer has yet to see widespread adoption in the active sonar community. There may be several reasons for this. For one, the method is not inherently robust, and will suffer from a phenomena called signal cancellation in active systems. Another reason is that in its original form, the computational complexity is cubic with the number of channels, O(M$^3$), while conventional beamformers are usually only at O($M$).

We will show that the suggested means for making the \gls{MVDR} behave well in active systems will also make it less computationally complex. Furthermore, by implementing it on a modern \gls{GPU} we have gained a speedup of around 2 orders of magnitude compared to an optimized single thread implementation in C.

@@

The receiver outputs can then be coherently combined in the post-processing to focus on specific points in the image. The combination process involves assigning suitable delays and weights to the array's channels, and is commonly referred to as beamforming.

@@

In a sonar processing chain, using one or more \glspl{GPU} can help offload the CPU.
